---
title: "Reproducing power analysis based on REACTIV validation studies of publication"
author: "Inka Marie Spyridonov"
date: "2024-12-17"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Power analysis based on REACTIV validation studies}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Reproducible environment

In case you want to make sure that you are creating the exact same output as presented in the publication, you need to create the reproducible environment using `rix` and `nix`. Using `rix`, you can run the following function to create your nix script. The nix script is however already in the vignettes folder within `xeredar` available as default.nix file.

```{r eval = FALSE}
rix::rix(
  r_ver = "4.4.0",
  r_pkgs = c("tidyverse", "tidyr", "purrr", "furrr", "gt","easystats","rlang", "DHARMa", "devtools", "patchwork", "rmarkdown", "ggpubr"),
  system_pkgs = NULL,
  tex_pkgs = c("amsmath"),
  git_pkgs = list(package_name = "xeredar",
                  repo_url = "https://github.com/basf/xeredar",
                  commit = "2b85763"),
  ide = "rstudio",
  project_path = "./vignettes",
  overwrite = TRUE
)

```

Once the default.nix file was created in the path you specified in the command above (`project_path`) you need to type `nix-build` into the terminal. Once the shell was built you can open it via `nix-shell`. You can open Rstudio by first running this line in the terminal: `export QT_XCB_GL_INTEGRATION=none` and then running `rstudio`. Rstudio will open and you will have access to all files located at the path you specified in the `rix` function. You can run the script as usual. Through this approach, the code and results are reproducible.

The following sections contain the code that needs to be executed to reproduce the analysis.

## Variance component analysis

### Functions

```{r eval=FALSE}
GetVarianceInfo <- function(ListAnovas, resultList){
  for(i in names(ListAnovas)){
    MeanControl <- insight::get_intercept(ListAnovas[[i]])
    VarianceControl <- insight::get_varcov(ListAnovas[[i]])[1,1]
    SDExperiment <- insight::get_sigma(ListAnovas[[i]])
    CVControl <- SDExperiment/MeanControl
    VarianceRep <- insight::get_variance(ListAnovas[[i]])$var.intercept[2]
    VarianceConcRep <- insight::get_variance(ListAnovas[[i]])$var.intercept[1]
    ICC <- performance::icc(ListAnovas[[i]], tolerance = 0)
    print(ICC)
    simultedResiduals <- DHARMa::simulateResiduals(ListAnovas[[i]], plot = F)
    zero_inflated <- as.numeric(DHARMa::testZeroInflation(simultedResiduals)[["p.value"]])
    if(!is.list(ICC)){
      ICC <- data.frame(adj_ICC = 0)
    } else{
      ICC <- ICC |> broom::tidy() 
      ICC <- ICC[1,3] |> rename(adj_ICC = mean)
    }
    resultList[[i]]$MeanControl <- MeanControl
    resultList[[i]]$SDExperiment <- SDExperiment
    resultList[[i]]$CVControl <- CVControl
    resultList[[i]]$SDRep <- sqrt(VarianceRep)
    resultList[[i]]$SDConcRep <- sqrt(VarianceConcRep)
    resultList[[i]]$ICC <- ICC
    resultList[[i]]$zeroInflated <- zero_inflated
    
  }
resultList
  }
```

### Analysis

```{r eval=TRUE, echo=TRUE, results='hide', message=FALSE}
devtools::install_github("thomasp85/patchwork")
#devtools::install_github("basf/xeredar")
purrr::map(c("dplyr","tidyr", "purrr", "furrr", "gt","easystats","rlang","drc", "DHARMa", "devtools", "patchwork", "ggplot2"), require, character.only = TRUE)
library(patchwork)
library(dplyr)
```

```{r eval=FALSE}

data <- xeredar::REACTIV_valid_data_table_spiked_unspiked |> 
  purrr::list_flatten()

```

```{r eval=FALSE}
resultList <- list()
ANOVAs <- furrr::future_map(data, ~ purrr::possibly(function(df) {
  lme4::lmer(Fluor ~ Conc + (1|Replicate) + (1|Replicate:Conc), data = df, REML = TRUE)
}, otherwise = "error", quiet = FALSE)(.x))

Variance_info <- list()
Variance_Info <- GetVarianceInfo(ListAnovas = ANOVAs, Variance_info) 

check <- purrr::imap(Variance_Info, ~ dplyr::tibble(MeanControl = .x$MeanControl,
  SDExperiment = .x$SDExperiment,
  CVControl = .x$CVControl, 
  SDRep = .x$SDRep,
  SDConcRep = .x$SDConcRep,
  adj_ICC = .x$ICC$adj_ICC,
  Spiked = !stringr::str_detect(.y, "Unspiked"),
  Compound = stringr::str_extract(.y, "^[^_]+"),
  Country = stringr::str_extract(.y, "(?<=_)[^_]+(?=_)"),
  zeroInflated = .x$zeroInflated))


Simulation_Info <- bind_rows(check)

SimulationInfoWTzeroInflated <- Simulation_Info  |> 
  filter(zeroInflated > 0.05) |> 
  dplyr::mutate(Inert = dplyr::if_else(
      Compound %in% c("Anastrozole", "anastrozole", "Amantadine", "Arabinose", "Atenolol", "Cefuroxime", "cefuroxime", "Cromolyn", "cromolyn", "Saccharin", "saccharin"), "Inert", "Active"
  )) |>
  dplyr::relocate(Compound, Inert) 

```

```{r eval=TRUE}



gt::gt(xeredar::SimulationInfoWTzeroInflated) |> gt::tab_caption(caption = "Summary table of information from the validation studies determined using the GetVarianceInfo() function. The first column specifies the respective compound tested, the second provides information whether it is an inert or an ective compound and the third column provides information about the estimated mean control measurement of the respective study. The fourth column refers to the standard deviation throughout the experiment, also referred to as residual standard deviation and the fifth column contains the Coefficient of Variation (CV) determined by dividing the SDExperiment column with the MeanControl column. The SDRep column contains information about the standard deviation explained through the replicate, followed by the standard deviation explained by the replicate treatment interaction. The adj_ICC column contains the adjusted intra-class Correlation Coefficient, followed by the column Spiked which specifies whether the previous data referes to spiked samples (TRUE) or unspiked samples (FALSE). The last column (Country) contains information about the lab in which the respective study was carried out.")
```

```{r}

  
result1 <- xeredar::SimulationInfoWTzeroInflated |> 
  dplyr::filter(Spiked == TRUE) |>
  rstatix::t_test(CVControl ~ Inert, var.equal = FALSE,
  alternative = "two.sided",) |>
  rstatix::add_significance() |> 
  rstatix::add_xy_position(x = "Inert") |> 
  mutate_if(is.numeric, ~round(., digits = 2))

xeredar::SimulationInfoWTzeroInflated |> 
  dplyr::filter(Spiked == TRUE) |>
  ggpubr::ggboxplot(x= "Inert", y= "CVControl", fill = "purple", size = NULL,
  width = 0.7, outlier.size = 0.3, outlier.shape = 16, lwd=0.1)+
  ggpubr::stat_pvalue_manual(result1, 
                             label = "Welch two-sample T-test, t= {statistic}, p = {p}", label.size = 1.5, bracket.size = 0.1)+
  ggplot2::stat_summary(fun="mean", color="darkred", size = 0.05)+
  xlab("Studies with spiked treatments")+
  ylab("Coefficient of Variation (CV)")+
   theme(axis.text = element_text(size=3),
         axis.title = element_text(size=4), 
         legend.text = element_text(size=2), 
         legend.title = element_text(size=3),
         axis.line = element_line(linewidth = 0.1), 
         axis.ticks = element_line(linewidth = 0.1))

#ggsave("CVSpiked.png", p1, width = 736, height = 518, units = "px")

result2 <- xeredar::SimulationInfoWTzeroInflated |> 
  dplyr::filter(Spiked == FALSE) |>
  rstatix::t_test(CVControl ~ Inert, var.equal = FALSE,
  alternative = "two.sided",) |>
  rstatix::add_significance() |> 
  rstatix::add_xy_position(x = "Inert") |> 
  mutate_if(is.numeric, ~round(., digits = 2))

xeredar::SimulationInfoWTzeroInflated |> 
  dplyr::filter(Spiked == FALSE) |>
  ggpubr::ggboxplot(x= "Inert", y= "CVControl", fill = "#00AFBB", size = NULL,
  width = 0.7, outlier.size = 0.3, outlier.shape = 16, lwd=0.1)+
  ggpubr::stat_pvalue_manual(result2, 
                             label = "Welch two-sample T-test, t= {statistic}, p = {p}", label.size = 1.5, bracket.size = 0.1)+
  ggplot2::stat_summary(fun="mean", color="darkred", size = 0.05)+
  xlab("Studies with unspiked treatments")+
  ylab("Coefficient of Variation (CV)")+
   theme(axis.text = element_text(size=3),
         axis.title = element_text(size=4), 
         legend.text = element_text(size=2), 
         legend.title = element_text(size=3),
         axis.line = element_line(linewidth = 0.1), 
         axis.ticks = element_line(linewidth = 0.1))
```

```{r eval=TRUE}
xeredar::SimulationInfoWTzeroInflated |> 
  dplyr::select(adj_ICC, Country, Spiked) |> 
  dplyr::mutate(Spiked = dplyr::if_else(Spiked == TRUE, "Spiked", "Unspiked")) |>
  ggplot2::ggplot(aes(y= as.numeric(adj_ICC), x= Country, color = Spiked)) +
  geom_jitter(alpha= 0.5,  width = 0.2, size = 0.3) +
  geom_abline(slope=0, intercept= 0.139, lwd= 0.1) +
  scale_color_manual(values = c("Spiked" = "purple", "Unspiked" = "#00AFBB")) + 
  labs(color='Spiked or Unspiked?') +
  ylab(label="adjusted intraclass Correlation Coefficient (ICC)") +
  xlab(label="Labs/country") +
  theme(panel.background = element_blank(),
        panel.grid.major = element_line(color = "lightgray", size = 0.1), 
        panel.grid.minor = element_line(color = "lightgray", size = 0.05),
        axis.text = element_text(size=3),
        axis.title = element_text(size=4), 
        legend.text = element_text(size=2), 
        legend.title = element_text(size=3),
        legend.key.size = unit(0.3, "lines"), 
        legend.spacing = unit(0.1, "lines"),
        axis.line = element_line(linewidth = 0.1), 
        axis.ticks = element_line(linewidth = 0.1)
        )
#ggsave("ICC.png", p3, width = 813, height = 566, units = "px")
```

## Data simulation

### Functions

```{r eval=FALSE}

# function for lognormal dose response model for simulating the increasing scenarios
LogNorm <- function(dose, b, c, d, e) {
            .expr1 <- d - c
            .expr4 <- log(dose) - log(e)
            .expr5 <- b * .expr4
            .expr6 <- pnorm(.expr5[1,1])
            .expr9 <- dnorm(.expr5[1,1])
            .value <- c + .expr1 * .expr6
            .value
            }

# function for generating the expected mean values for each simulated concentration based on information of the original studies and the respective simulated inflection point of the log-normal model 
createMeans <- function(dfEmpty, variabledf,c = 2.5e+05,b=1.2){
  for(i in 1:nrow(variabledf)){
  y <- LogNorm(dose = variabledf[i,"Conc"], b = b, c = c , d = c*4, e = variabledf[i,"e"])
  dfEmpty[[i]] <- data.frame("e" = variabledf[i,"e"], "Conc" = variabledf[i,"Conc"], "Mean" = y, "percEffect" = ((y-c)/c)*100)
  }
  dfEmpty
}

# function for 
MeanValuesAllStudies <- function(SimulationInfo, dfEmpty, variabledf, b = 1.2){
  MeanValuesStudy <- list()
  for(i in 1:nrow(SimulationInfo)){
    Means <- dplyr::bind_rows(createMeans(dfEmpty = dfEmpty, variabledf = variabledf, c = as.numeric(SimulationInfo[[i, "MeanControl"]]), b = b))
    Means$ValidStudy <- paste(Means$e, i)
    Means$MeanControl <- SimulationInfo[[i, "MeanControl"]]
    Means$SDExperiment <- SimulationInfo[[i, "SDExperiment"]]
    Means$CVControl <- SimulationInfo[[i, "CVControl"]]
    Means$SDRep <- SimulationInfo[[i,"SDRep"]]
    Means$SDConcRep <- SimulationInfo[[i,"SDConcRep"]]
    MeanValuesStudy[[i]] <- Means
  }
  MeanValuesStudy
}


SimulateNow_HomoVar <- function(Spiked = TRUE, MeanValues, nEmbryoperConc=8, nConc=6, nRep=3, doses=c(0,0.16,0.8,4,20,100), Repetitions=1000)
  {
  simulatedDF <- list()
  simulatedList <- list()
  for(r in 1:Repetitions){
    
    if(Spiked==TRUE){
      
      for(i in MeanValues[,"ValidStudy"]){
        
        StudySelect <- MeanValues |> filter(ValidStudy==i)
        StudySelect$sdPredMean <- as.numeric(StudySelect$SDExperiment)
        dosesSp <- c(c(0),doses)
        df <- data.frame(Replicate =rep(1:nRep, each=(nEmbryoperConc*nConc)+nEmbryoperConc), Conc = rep(dosesSp, each=nEmbryoperConc, times=3))
        df$RepRand <- rep(rnorm(max(df$Replicate),0, as.numeric(unique(StudySelect$SDRep))),each=(nEmbryoperConc*nConc)+nEmbryoperConc)
        ConcRepRand <- rnorm(nrow(unique(df[c("Conc","Replicate")])),0, as.numeric(unique(StudySelect$SDConcRep)))
        df$ConcRepRand <- rep(c(ConcRepRand[1], ConcRepRand[1:6],ConcRepRand[7], ConcRepRand[7:12], ConcRepRand[13], ConcRepRand[13:18]), each=nEmbryoperConc)
        df$Mean <- left_join(df,StudySelect, by=join_by(Conc))[,"Mean"]
        df$FluorFixed <- purrr::modify2(df$Mean, left_join(df,StudySelect, by=join_by(Conc))[,"sdPredMean"], ~ rnorm(n = 1,mean = .x, sd=.y))
        df$Fluor <- df$FluorFixed+df$RepRand+df$ConcRepRand
        simulatedDF[[i]] <- df[,c("Replicate","Conc","Fluor")]
        }
      }
    else{
      
      for(i in MeanValues[,"ValidStudy"]){
        
        StudySelect <- MeanValues |> filter(ValidStudy==i)
        StudySelect$sdPredMean <- as.numeric(StudySelect$SDExperiment)
        df <- data.frame(Replicate =rep(1:nRep, each=nEmbryoperConc*nConc), Conc = rep(doses, each=nEmbryoperConc, times=3))
        df$RepRand <- rep(rnorm(max(df$Replicate),0, as.numeric(unique(StudySelect$SDRep))),each=nEmbryoperConc*nConc)
        df$ConcRepRand <- rnorm(nrow(unique(df[c("Conc","Replicate")])),0, as.numeric(unique(StudySelect$SDConcRep)))
        df$Mean <- left_join(df,StudySelect, by=join_by(Conc))[,"Mean"]
        df$FluorFixed <- purrr::modify2(df$Mean, left_join(df,StudySelect, by=join_by(Conc))[,"sdPredMean"], ~ rnorm(n = 1,mean = .x, sd=.y))
        df$Fluor <- df$FluorFixed+df$RepRand+df$ConcRepRand
        df[,c("Replicate","Conc","Fluor")]
        simulatedDF[[i]] <- df[,c("Replicate","Conc","Fluor")]
      }
      
    }
    
    simulatedList[[r]] <- simulatedDF
  }
  simulatedList
  
}
```

```{r eval=FALSE}

Conc <- c(0,0.16,0.8,4,20,100)
e <- c(9,20, 49,81, 123, 153,200,260,330,400, Inf)

variables <- expand_grid(e = e,Conc = Conc)

DFwithMeansperConc <- list()
SimInfoSpiked <- xeredar::SimulationInfoWTzeroInflated |> dplyr::filter(Spiked == TRUE)
SimInfoUnspiked <- xeredar::SimulationInfoWTzeroInflated |> dplyr::filter(Spiked == FALSE)

dfEmpty <- list()
MeansAllStudiesperConcSpiked <- dplyr::bind_rows(MeanValuesAllStudies(SimulationInfo = SimInfoSpiked, dfEmpty = dfEmpty,variabledf = variables, b = 1.2 ))

MeansAllStudiesperConcUnspiked <- dplyr::bind_rows(MeanValuesAllStudies(SimulationInfo = SimInfoUnspiked, dfEmpty = dfEmpty,variabledf = variables, b = 1.2 ))

SimDatSpiked_Homo <-  SimulateNow_HomoVar(Spiked=TRUE, MeanValues = MeansAllStudiesperConcSpiked, Repetitions=25)

SimDatUnspiked_Homo <- SimulateNow_HomoVar(Spiked=FALSE, MeanValues = MeansAllStudiesperConcUnspiked, Repetitions=25)
```

## Analyzing simulated data

```{r eval=FALSE}

ConvertStructure <- function(ListObject){
  for(i in 1:length(ListObject)){
  ListObject[[i]]$Conc <- factor(ListObject[[i]]$Conc, ordered=TRUE, levels = sort(unique(ListObject[[i]]$Conc)))
  ListObject[[i]]$Replicate <- factor(ListObject[[i]]$Replicate)
  ListObject[[i]]$Treatment <- ListObject[[i]]$Conc
  }
  ListObject
}
SimDatSpikedFLAT <- purrr::list_flatten(SimDatSpiked_Homo, name_repair="universal")
SimDatUnspikedFLAT <- purrr::list_flatten(SimDatUnspiked_Homo, name_repair="universal")

SimDatSpikedFLAT<- ConvertStructure(ListObject=SimDatSpikedFLAT)
SimDatUnspikedFLAT <- ConvertStructure(ListObject=SimDatUnspikedFLAT)

ResultsSpiked_dataPrep_alpha0 <- SimDatSpikedFLAT |> furrr::future_map(\(df) xeredar::data_prep(df, row = FALSE, trimming = FALSE, boxcox = FALSE, alpha = 0)[c("Dunnetts", "WilliamsDecrease", "WilliamsIncrease", "Justify", "Monotonicity Test")])
usethis::use_data(ResultsSpiked_dataPrep_alpha0)

ResultsUnspiked_dataPrep_alpha0 <- SimDatUnspikedFLAT |> furrr::future_map(\(df) xeredar::data_prep(df, row = FALSE, trimming = FALSE, boxcox = FALSE, alpha = 0)[c("Dunnetts", "WilliamsDecrease", "WilliamsIncrease", "Justify", "Monotonicity Test")])
usethis::use_data(ResultsUnspiked_dataPrep_alpha0)


MeanSpikedControls <- function(ListObject){
  ResultList <- list()
  for(i in 1:length(ListObject)){
    ControlMeans <- ListObject[[i]] |> 
      filter(Conc=="0") |> 
      summarise(mean(Fluor, na.rm=TRUE), .by=Replicate)
    ResultList[[i]] <- ControlMeans
  }
  ResultList
}
NormalizationFunc <- function(MeanDataList, SimuDataList){
  resultList <- list()
  for(i in 1:length(SimuDataList)){
    SimuDat <- SimuDataList[[i]]
    MeanDat <- MeanDataList[[i]]
    SimuDat$NormFluor <- NA
    for(k in 1:nrow(SimuDat)){
      
      if(SimuDat[k,"Replicate"]=="1"){
        
      SimuDat[k,"NormFluor"] <- SimuDat[k,"Fluor"]/MeanDat[1,2]
        }else if(SimuDat[k,"Replicate"]=="2"){
      SimuDat[k,"NormFluor"] <- SimuDat[k,"Fluor"]/MeanDat[2,2]
      }else{
      SimuDat[k,"NormFluor"] <- SimuDat[k,"Fluor"]/MeanDat[3,2]
      }
      
    }
      
      
    resultList[[i]] <- SimuDat
  }
  resultList
}
SimInfoSpiked <- Simulation_Info |> dplyr::filter(Spiked == TRUE)
SimInfoUnspiked <- Simulation_Info |> dplyr::filter(Spiked == FALSE)

SimInfoSpiked <- rowid_as_column(SimInfoSpiked)
SimInfoUnspiked <- rowid_as_column(SimInfoUnspiked)

SimInfoMatch_SpikedUnspiked <- full_join(SimInfoSpiked |> dplyr::select(rowid, Country, Compound),
          SimInfoUnspiked |> dplyr::select(rowid, Country, Compound), by= c("Compound", "Country")
          ) |> rename(id_spiked = rowid.x, id_unspiked = rowid.y)
MeanSpikedControlsRep <- MeanSpikedControls(SimDatSpikedFLAT)

SimInfoMatch_SpikedUnspiked <- SimInfoMatch_SpikedUnspiked |> filter(!is.na(id_unspiked))

SimDatSpikedFLAT_ForUnspiked <- SimDatSpikedFLAT 

names(SimDatSpikedFLAT_ForUnspiked) <-
  stringr::str_extract(names(SimDatSpikedFLAT), pattern = "(?<=\\.\\.\\d{1,3}\\.|Inf\\.)\\d+")

SimDatSpikedFLAT_ForUnspiked <- SimDatSpikedFLAT_ForUnspiked[names(SimDatSpikedFLAT_ForUnspiked) %in% as.character(SimInfoMatch_SpikedUnspiked$id_unspiked)]

MeanSpikedControlsRep_UNSPIKED <- MeanSpikedControls(SimDatSpikedFLAT_ForUnspiked)


SimDatUnspikedFLATNormalized <- NormalizationFunc(MeanSpikedControlsRep_UNSPIKED, SimDatUnspikedFLAT)

SimDatSpikedFLATNormalized <- NormalizationFunc(MeanSpikedControlsRep, SimDatSpikedFLAT)
possDunneTT = possibly(.f = PMCMRplus::dunnettTest, otherwise = "Error")


ResultsUnspiked_pooledDunnett  <- SimDatUnspikedFLATNormalized |> furrr::future_map(\(df) possDunneTT(df$NormFluor,df$Conc, alternative = "two.sided"), .options = furrr::furrr_options(seed = TRUE))
usethis::use_data(ResultsUnspiked_pooledDunnett)
names(ResultsUnspiked_pooledDunnett) <- names(ResultsUnspiked_dataPrep_alpha0)
usethis::use_data(ResultsUnspiked_pooledDunnett, overwrite = TRUE)

ResultsSpiked_pooledDunnett  <- SimDatSpikedFLATNormalized |> furrr::future_map(\(df) possDunneTT(df$NormFluor,df$Conc, alternative = "two.sided"), .options = furrr::furrr_options(seed = TRUE))
names(ResultsSpiked_pooledDunnett) <- names(ResultsSpiked_dataPrep_alpha0)
usethis::use_data(ResultsSpiked_pooledDunnett, overwrite = TRUE)

possDunns = possibly(.f = PMCMRplus::kwManyOneDunnTest, otherwise = "Error")


ResultsUnspiked_pooledDunns <- SimDatUnspikedFLATNormalized |> furrr::future_map(\(df) possDunns(df$NormFluor,df$Conc, alternative = "two.sided", p.adjust.method = "holm"), .options = furrr::furrr_options(seed = TRUE))
names(ResultsUnspiked_pooledDunns) <- names(ResultsUnspiked_dataPrep_alpha0)
usethis::use_data(ResultsUnspiked_pooledDunns, overwrite = TRUE)

ResultsSpiked_pooledDunns  <- SimDatSpikedFLATNormalized |> furrr::future_map(\(df) possDunns(df$NormFluor,df$Conc, alternative = "two.sided", p.adjust.method = "holm"), .options = furrr::furrr_options(seed = TRUE))
names(ResultsSpiked_pooledDunns) <- names(ResultsSpiked_dataPrep_alpha0)
usethis::use_data(ResultsSpiked_pooledDunns, overwrite = TRUE)
```

## Presenting results

```{r eval=FALSE}

extract_interestMixed <- function(input_list, name="WilliamsIncrease", name_pval="PvalmixedWilliamsIncrease", test_type= "williams"){
  result <-  purrr::list_flatten(input_list)
  result <- result[names(result)[stringr::str_detect(names(result),name)]]
  names(result) <- str_remove(names(result), paste0("_",name))
  result <- purrr::list_flatten(result, name_repair = "universal")
  result <- dplyr::bind_rows(result, .id = "StudyID")
  if(test_type == "williams"){
    result <- mutate(result, !!name_pval := ifelse(result$Sign == TRUE, 1, 0))
    result$Conc = as.numeric(result$Conc)
  } else if( test_type == "dunnetts_mixed"){
    result <- mutate(result, !!name_pval := ifelse(result$`adj p` < 0.05, 1, 0))
    result$Conc <- as.numeric(stringr::str_extract(row.names(result), "\\d+.\\d+|\\d+"))
  } 
  result
}


extract_interestPooled <- function(input_list, name_pval="SignPooledDunnetts"){
  
  result <-  purrr::list_flatten(input_list, name_repair="universal")
  
  possPMCMRTable = possibly(.f = rcompanion::PMCMRTable, otherwise = "Error")
    
  result <- input_list |> furrr::future_map(\(df) possPMCMRTable(df,reverse = TRUE,   digits = 3))
  result <- dplyr::bind_rows(result, .id="StudyID")
  result$Conc <- as.numeric(stringr::str_extract(result$Comparison, "^\\d+.\\d+|^\\d+"))
  result <- mutate(result, !!name_pval := ifelse(as.numeric(result$p.value) < 0.01, 1, 0))
}

get_results_table <- function(EC50_values = c("\\.\\.9\\.", "\\.\\.20\\.", "\\.\\.49\\.","\\.\\.81\\.", "\\.\\.123\\.", "\\.\\.153\\.",  "\\.\\.200\\.", "\\.\\.260\\.", "\\.\\.330\\.", "\\.\\.400\\.", "Inf\\."), Spiked = TRUE){
  
  furrr::future_map(EC50_values, function(x){
    
    spike_type <- ifelse(Spiked, "Spiked", "Unspiked")
    Name <- names(get(paste0("Results", spike_type, "_pooledDunnett")))[stringr::str_starts(names(get(paste0("Results", spike_type, "_pooledDunnett"))),x)]

    # Assuming the required data is in your global environment
    # Replace with actual names of your data
    results_Dunnetts <- get(paste0("Results", spike_type, "_pooledDunnett"))[Name]
    PooledDunnetts <- extract_interestPooled(input_list = results_Dunnetts, name_pval = "SignPooledDunnetts")
    
    results_Dunns <- get(paste0("Results", spike_type, "_pooledDunns"))[Name]
    pooledDunns <- extract_interestPooled(input_list = results_Dunns, name_pval = "SignpooledDunns")
    
    results_dataPrep_alpha0 <- get(paste0("Results", spike_type, "_dataPrep_alpha0"))[Name]
    mixedDunnett <- extract_interestMixed(input_list= results_dataPrep_alpha0, name = "Dunnetts", name_pval = "PvalmixedDunnett", test_type = "dunnetts_mixed" )
    mixedWilliamsIncrease <- extract_interestMixed(input_list= results_dataPrep_alpha0, name = "WilliamsIncrease", name_pval = "PvalmixedWilliamsIncrease", test_type = "williams" )
    
    dfs = list(
      mixedDunnett[,c("Conc","StudyID","PvalmixedDunnett")],
      PooledDunnetts[,c("Conc","StudyID","SignPooledDunnetts")],
      pooledDunns[,c("Conc","StudyID","SignpooledDunns")],
      mixedWilliamsIncrease[,c("Conc","StudyID","PvalmixedWilliamsIncrease")]
    )
    
    df <- purrr::reduce(dfs, dplyr::full_join, by=c("Conc","StudyID"))
    
    result <- df %>% dplyr::summarise(
      mixedDunnett = mean(PvalmixedDunnett, na.rm=TRUE), 
      PooledDunnetts = mean(SignPooledDunnetts, na.rm=TRUE),
      pooledDunns = mean(SignpooledDunns, na.rm = TRUE),
      mixedWilliamsIncrease = mean(PvalmixedWilliamsIncrease, na.rm = TRUE),
      .by = Conc
    )  %>% 
      tidyr::pivot_longer(cols = mixedDunnett:mixedWilliamsIncrease, names_to = "Test", values_to = "PercentSign") 

    result
  })
}
```

```{r eval=FALSE}

resultsSpiked_power_reactive  <- get_results_table()
resultsUnspiked_power_reactive <- get_results_table(Spiked = FALSE)

resultsSpiked_power_reactive <- xeredar::resultsSpiked_power_reactive

names(resultsSpiked_power_reactive) <- c("9", "20", "49", "81", "123", "153", "200", "260", "330", "400", "Inf")

perc_effect_e_dose_Spiked <- xeredar::MeansAllStudiesperConcSpiked |> 
  dplyr::select(e, Conc, percEffect) |> 
  dplyr::mutate(across(everything(), round, digits = 2 )) |>
  dplyr::distinct() |>
  dplyr::mutate(e = as.character(e))

resultsSpiked_power_reactive  <- dplyr::bind_rows(resultsSpiked_power_reactive, .id = "e") 

resultsSpiked_power_reactive <- dplyr::left_join(resultsSpiked_power_reactive, perc_effect_e_dose_Spiked, by = c("e", "Conc"))

resultsUnspiked_power_reactive <- xeredar::resultsUnspiked_power_reactive

names(resultsUnspiked_power_reactive) <- c("9", "20", "49", "81", "123", "153", "200", "260", "330", "400", "Inf")

perc_effect_e_dose_Unspiked <- xeredar::MeansAllStudiesperConcUnspiked |> 
  dplyr::select(e, Conc, percEffect) |> 
  dplyr::mutate(across(everything(), round, digits = 2 )) |>
  dplyr::distinct() |>
  dplyr::mutate(e = as.character(e))


resultsUnspiked_power_reactive  <- dplyr::bind_rows(resultsUnspiked_power_reactive, .id = "e") 

resultsUnspiked_power_reactive <- dplyr::left_join(resultsUnspiked_power_reactive, perc_effect_e_dose_Unspiked)

usethis::use_data(resultsUnspiked_power_reactive, overwrite = TRUE)
```

```{r eval=TRUE, warning=FALSE}
breaks=c("pooledDunns","PooledDunnetts","mixedWilliamsIncrease", "mixedDunnett")

labels=c("Pooled Dunns","Pooled Dunnetts","mixed Williams","mixed Dunnett")

resultsSpiked_power_reactive <- xeredar::resultsSpiked_power_reactive

type1_error_Spiked <- resultsSpiked_power_reactive |> 
  dplyr::arrange(e, percEffect) |> 
  dplyr::filter(percEffect <= 1) |>
  ggplot2::ggplot(ggplot2::aes(x=PercentSign, color=Test))+
  ggplot2::geom_boxplot(alpha = 0.5, lwd=0.1, outlier.size = 0.15)+
  ggplot2::scale_x_continuous(labels = scales::percent, trans = "log10")+
  ggplot2::scale_color_discrete(name  ="Tests",
                      breaks = breaks,
                      labels = labels) + 
  ggplot2::xlab("% significant tests") +
  ggplot2::ylab("") +
  theme(panel.background = element_blank(),
        panel.grid.major = element_line(color = "lightgray", size = 0.1), 
        panel.grid.minor = element_line(color = "lightgray", size = 0.05),
        axis.text = element_text(size=3),
        axis.title = element_text(size=4), 
        legend.text = element_text(size=2), 
        legend.title = element_text(size=3),
        legend.key.size = unit(0.3, "lines"), 
        legend.spacing = unit(0.1, "lines"),
        axis.line = element_line(linewidth = 0.1), 
        axis.ticks = element_line(linewidth = 0.1),
        axis.ticks.y = element_blank(),
        axis.text.y=element_blank()
        )

#ggsave("type1_error_Spiked.png", type1_error_Spiked, width = 736, height = 518, units = "px")


breaks=c("mixedWilliamsIncrease", "mixedDunnett","pooledDunns","PooledDunnetts")

labels=c("mixed Williams","mixed Dunnett","Pooled Dunns","Pooled Dunnetts")

power_Spiked <- resultsSpiked_power_reactive |> 
  dplyr::arrange(e, percEffect) |>  
  dplyr::filter(Conc==100) |>
  ggplot2::ggplot(ggplot2::aes(x=percEffect, y=PercentSign, color=Test))+
  ggplot2::geom_line(linewidth=0.2)+
  ggplot2::geom_hline(x=as.numeric(0:300), yintercept=0.8, lwd = 0.2)+
  ggplot2::scale_color_manual(name  ="Approach",
                      breaks = breaks,
                      labels = labels,
                      values=c("#00e193","#312e9c","#d8d343","#920011"))+
  ggplot2::ylab("% Significant test results \n in the highest tested concentration")+
  ggplot2::xlab("Simulated % effect") + 
  theme(panel.background = element_blank(),
        panel.grid.major = element_line(color = "lightgray", size = 0.1), 
        panel.grid.minor = element_line(color = "lightgray", size = 0.05),
        axis.text = element_text(size=3),
        axis.title = element_text(size=4), 
        legend.text = element_text(size=2), 
        legend.title = element_text(size=3),
        legend.key.size = unit(0.3, "lines"), 
        legend.spacing = unit(0.1, "lines"),
        axis.line = element_line(linewidth = 0.1), 
        axis.ticks = element_line(linewidth = 0.1)
        )
#ggsave("power_Spiked.png", power_Spiked, width = 736, height = 518, units = "px")


power_Unspiked <- xeredar::resultsUnspiked_power_reactive |> 
  dplyr::arrange(e, percEffect) |>  
  dplyr::filter(Conc==100) |>
  ggplot2::ggplot(ggplot2::aes(x=percEffect, y=PercentSign, color=Test))+
  ggplot2::geom_line(linewidth=0.2)+
  ggplot2::geom_hline(x=as.numeric(0:300), yintercept=0.8, lwd = 0.2)+
  ggplot2::scale_color_manual(name  ="Approach",
                      breaks = breaks,
                      labels = labels,
                      values=c("#00e193","#312e9c","#d8d343","#920011"))+
  ggplot2::ylab("% Significant test results \n in the highest tested concentration")+
  ggplot2::xlab("Simulated % effect") + 
  theme(panel.background = element_blank(),
        panel.grid.major = element_line(color = "lightgray", size = 0.1), 
        panel.grid.minor = element_line(color = "lightgray", size = 0.05),
        axis.text = element_text(size=3),
        axis.title = element_text(size=4), 
        legend.text = element_text(size=2), 
        legend.title = element_text(size=3),
        legend.key.size = unit(0.3, "lines"), 
        legend.spacing = unit(0.1, "lines"),
        axis.line = element_line(linewidth = 0.1), 
        axis.ticks = element_line(linewidth = 0.1)
        )
#ggsave("power_Unspiked.png", power_Unspiked, width = 736, height = 518, units = "px")



power_Spiked / power_Unspiked / type1_error_Spiked
```