---
title: "Reproducing power analysis of publication"
author: "Inka Marie Spyridonov"
date: "2024-12-17"
vignette: >
  %\VignetteIndexEntry{User guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


To reproduce the power analysis in the publication the following code has to be executed: 


---
title: "ValidationStudyDataExploration"
author: "Inka Spyridonov"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

## Functions

```{r}
#| label: Settings
#| include: false
rm(list = ls())
#devtools::install_git("https://gitlab.roqs.basf.net/WillmsIM/xetaana.git")
#install.packages("easystats")
lapply(c("readxl","tidyverse","xeredar","ARTool","broom","car","dplyr","emmeans","ggplot2","lme4","magrittr","MASS","multcomp","psych","tidyr", "kableExtra","knitr","purrr","gt","easystats","rlang","drc","truncnorm", "DHARMa", "devtools"), require, character.only=TRUE)

```



```{r}

data <- xeredar::REACTIV_valid_data_table_spiked_unspiked |> 
  purrr::list_flatten()

```


```{r}
#| label: FunctionGetVarianceInfo
#| include: false
GetVarianceInfo <- function(ListAnovas, resultList){
  for(i in names(ListAnovas)){
    MeanControl <- insight::get_intercept(ListAnovas[[i]])
    VarianceControl <- insight::get_varcov(ListAnovas[[i]])[1,1]
    SDExperiment <- insight::get_sigma(ListAnovas[[i]])
    CVControl <- SDExperiment/MeanControl
    VarianceRep <- insight::get_variance(ListAnovas[[i]])$var.intercept[2]
    VarianceConcRep <- insight::get_variance(ListAnovas[[i]])$var.intercept[1]
    ICC <- performance::icc(ListAnovas[[i]], tolerance = 0)
    print(ICC)
    simultedResiduals <- DHARMa::simulateResiduals(ListAnovas[[i]], plot = F)
    zero_inflated <- as.numeric(DHARMa::testZeroInflation(simultedResiduals)[["p.value"]])
    if(!is.list(ICC)){
      ICC <- data.frame(adj_ICC = 0)
    } else{
      ICC <- ICC |> broom::tidy() 
      ICC <- ICC[1,3] |> rename(adj_ICC = mean)
    }
    resultList[[i]]$MeanControl <- MeanControl
    resultList[[i]]$SDExperiment <- SDExperiment
    resultList[[i]]$CVControl <- CVControl
    resultList[[i]]$SDRep <- sqrt(VarianceRep)
    resultList[[i]]$SDConcRep <- sqrt(VarianceConcRep)
    resultList[[i]]$ICC <- ICC
    resultList[[i]]$zeroInflated <- zero_inflated
    
  }
resultList
  }
```

### USA

#### Spiked

```{r}
#| include: false
resultList <- list()
ANOVAs <- furrr::future_map(data, ~ purrr::possibly(function(df) {
  lme4::lmer(Fluor ~ Conc + (1|Replicate) + (1|Replicate:Conc), data = df, REML = TRUE)
}, otherwise = "error", quiet = FALSE)(.x))

Variance_info <- list()
Variance_Info <- GetVarianceInfo(ListAnovas = ANOVAs, Variance_info)

```

```{r}
#| include: false

InfoUSAdfSpiked <-  dplyr::bind_rows(GetVarianceInfo("Spiked", resultUSAPerCompound, resultList = resultList), .id="Compound") |> mutate(Spiked = TRUE, Country = "USA")
```

#### Unspiked

```{r}
#| include: false
InfoUSAdfUnspiked <- dplyr::bind_rows(GetVarianceInfo("Unspiked", resultUSAPerCompound, resultList = resultList),.id="Compound") |> mutate(Spiked = FALSE, Country = "USA")
```

### UK

#### Spiked

```{r}
#| include: false
resultList <- list()
resultUK <- check(df= DataUK, resultList=resultList)

```

```{r}
#| include: false
InfoUKdfSpiked <- dplyr::bind_rows(GetVarianceInfo("Spiked", resultUK, resultList = resultList), .id="Compound") |> mutate(Spiked = TRUE, Country = "UK")
```

#### Unspiked

```{r}
#| include: false
InfoUKdfUnspiked <- dplyr::bind_rows(GetVarianceInfo("Unspiked", resultUK, resultList = resultList), .id="Compound") |> mutate(Spiked = FALSE, Country = "UK")
```

### Japan

#### Spiked

```{r}
#| include: false
resultList <- list()
resultJapan <- check(df= DataJapan, resultList=resultList)

```

```{r}
#| include: false
InfoJapandfSpiked <- dplyr::bind_rows(GetVarianceInfo("Spiked", resultJapan, resultList = resultList), .id="Compound") |> mutate(Spiked = TRUE, Country = "Japan")
```

#### Unspiked

```{r}
#| include: false
InfoJapandfUnspiked <- dplyr::bind_rows(GetVarianceInfo("Unspiked", resultJapan, resultList = resultList), .id="Compound") |> mutate(Spiked = FALSE, Country = "Japan")
```

### Germany

#### Spiked

```{r}
#| include: false
resultList <- list()
resultGermany <- check(df= DataGermany, resultList=resultList)

```

```{r}
#| include: false
InfoGermanydfSpiked <- dplyr::bind_rows(GetVarianceInfo("Spiked", resultGermany, resultList = resultList), .id="Compound") |> mutate(Spiked = TRUE, Country = "Germany")
```

#### Unspiked

```{r}
#| include: false
InfoGermanydfUnspiked <- dplyr::bind_rows(GetVarianceInfo("Unspiked", resultGermany, resultList = resultList), .id="Compound") |> mutate(Spiked = FALSE, Country = "Germany")
```

### France

#### Spiked

```{r}
#| include: false
resultList <- list()
resultFrance <- check(df= DataFrance, resultList=resultList)

```

```{r}
#| include: false
InfoFrancedfSpiked <- dplyr::bind_rows(GetVarianceInfo("Spiked", resultFrance, resultList = resultList), .id="Compound") |> mutate(Spiked = TRUE, Country = "France")
```

#### Unspiked

```{r}
#| include: false
InfoFrancedfUnspiked <- dplyr::bind_rows(GetVarianceInfo("Unspiked", resultFrance, resultList = resultList), .id="Compound") |> mutate(Spiked = FALSE, Country = "France")
```

### Denmark

#### Spiked

```{r}
#| include: false
#| eval: false
resultList <- list()
resultDenmark <- check(df= DataDenmark, resultList=resultList)

```

```{r}
#| eval: false
#| include: false
InfoDenmarkdfSpiked <- dplyr::bind_rows(GetVarianceInfo("Spiked", resultDenmark, resultList = resultList), .id="Compound") |> mutate(Spiked = TRUE, Country = "Denmark")
```

#### Unspiked

```{r}
#| include: false
#| eval: false
InfoDenmarkdfUnspiked <- dplyr::bind_rows(GetVarianceInfo("Unspiked", resultDenmark, resultList = resultList), .id="Compound") |> mutate(Spiked = FALSE, Country = "Denmark")

```

## Combining all countries

```{r}
SimulationInfo <- rbind(InfoUSAdfSpiked,InfoUSAdfUnspiked, InfoUKdfSpiked, InfoUKdfUnspiked, InfoJapandfSpiked, InfoJapandfUnspiked, InfoGermanydfSpiked, InfoGermanydfUnspiked, InfoFrancedfSpiked, InfoFrancedfUnspiked, InfoDenmarkdfSpiked, InfoDenmarkdfUnspiked)

```

### 

```{r}
#| echo: false
#| label: tbl_summaryInfoAll
#| tbl-cap: "Control mean and variance information as well as random term variance information from all labs and compounds. Furthermore, the spacing factor used in each study and the number of treatments is shown. This table serves as basis for data simulation."

SimulationInfo <- SimulationInfo |> unnest(cols=ICC) 

SimulationInfo <- SimulationInfo  |> filter(zeroInflated > 0.05)


save(SimulationInfo, file="SimulationInfoWTzeroInflated.rda")



```

## Justifying mixed Model

```{r}
SimulationInfo |> 
  dplyr::select(adj_ICC, Country, Spiked) |> 
  dplyr::mutate(Spiked = if_else(Spiked == TRUE, "Spiked", "Unspiked")) |>
  ggplot(aes(y= as.numeric(adj_ICC), x= Country, color = Spiked))+
  geom_jitter(alpha= 0.5,  width = 0.2)+
  geom_abline(slope=0, intercept= 0.1337804)+
  labs(color='Spiked or Unspiked?')+
  ylab(label="adjusted ICC value")+
  xlab(label="")
```
