---
title: "Analysis of all RADAR validation studies"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{User guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


## Reproducible environment

In case you want to make sure that you are creating the exact same output as presented in the publication, you need to create the reproducible environment using `rix` and `nix`. Using `rix`, you can run the following function to create your nix shell. 

```{r eval = FALSE}
rix::rix(
  r_ver = "latest",
  r_pkgs = c("tidyverse", "tidyr", "purrr", "furrr", "gt","easystats","rlang", "DHARMa", "devtools", "patchwork", "rmarkdown"),
  system_pkgs = NULL,
  tex_pkgs = c("amsmath"),
  git_pkgs = list(package_name = "xeredar",
                  repo_url = "https://github.com/basf/xeredar",
                  commit = "2b85763"),
  ide = "rstudio",
  project_path = "./vignettes",
  overwrite = TRUE
)

```

Once the default.nix file was created in the path you specified in the command above (`project_path`) you need to type `nix-build` into the terminal. Once the shell was built you can open it via `nix-shell`. You can open Rstudio by first running this line in the terminal: `export QT_XCB_GL_INTEGRATION=none` and then running `rstudio`. Rstudio will open and you will have access to all files located at the path you specified in the `rix` function. You can run the script as usual. Through this approach, the code and results are reproducible.

The following sections contain the code that needs to be executed to reproduce the analysis.


```{r}



# for(i in unique(RADAR_valid_data_table$Compound)){
#   for(j in unique(RADAR_valid_data_table$lab)){
#     dat <- RADAR_valid_data_table |> 
#       dplyr::filter(Compound == i & lab == j) 
#     if(!nrow(dat) == 0){
#       RADAR_valid_data[[paste0(i,j)]] <- dat
#     } else{
#       next
#     }
#   }
# }


decreasing_alpha <- function(df, vector){
  for(i in vector){
    result <- xeredar::data_prep(df, row=FALSE, alpha = i, trim=FALSE, outlier = FALSE)
    if(is.list(result)){
      break
    } else{
      next
    }
  }
  result[["alpha"]] <- i
  return(result)
}


# RADAR_valid_data_table_spiked_unspiked <- purrr::map(RADAR_valid_data, spiked_unspiked) 


changing_alpha <- sort(c(c(1,5) %o% 10^-(2:17), 0), decreasing = TRUE)[-1]


check <- function(list_dfs, id){
  result <- tryCatch({
    decreasing_alpha(list_dfs, changing_alpha)
    }, error = function(e) {
      list(error = paste("Error xeredar provides an error"))
      })
  setNames(list(result), id)
}


anovas <- purrr::imap(xeredar::RADAR_valid_data_table_spiked_unspiked, ~ check(.x, .y)) |> purrr::flatten()




check_positive <- function(df){
  if(length(df) ==1){
    return(NA)}
  
  if(df[["Monotonicity Test"]]$`Pr(>|t|)`[1] >= 0.01 &
     df[["Monotonicity Test"]]$`Pr(>|t|)`[2] <= 0.01 ) {
    return(df[["Dunnetts"]])
  } else if (length(unique(df$WilliamsIncrease$Y.Tilde)) - 1 > floor(nrow(df$WilliamsIncrease)/2) | 
             length(unique(df$WilliamsDecrease$Y.Tilde)) - 1 > floor(nrow(df$WilliamsDecrease)/2)) { 
    return(list(WilliamsIncrease = df[["WilliamsIncrease"]], WilliamsDecrease = df[["WilliamsDecrease"]])) 
  } else{
    return(df[["Dunnetts"]])
  } 
}


resultList <- 

anovas_unspiked <- anovas[stringr::str_detect(names(anovas), "Unspiked")]
anovas_spiked<- anovas[!stringr::str_detect(names(anovas), "Unspiked")]


results_Spiked <- purrr::map(anovas_spiked, check_positive)
results_Unspiked <- purrr::map(anovas_unspiked, check_positive)


Williams_Spiked <- !purrr::map_vec(results_Spiked, is.data.frame)
Williams_Unspiked <- !purrr::map_vec(results_Unspiked, is.data.frame)

significances <- function(df, Williams= FALSE){
  if(!is.list(df)){
    return(NA)}
  if(Williams) {
    if(TRUE %in% df[["WilliamsIncrease"]]$Sign){
        return("sign_Incr")
      }
    if(TRUE %in% df[["WilliamsDecrease"]]$Sign){
        return("sign_Decr")
      }
    else{
      return("Nonsignificant")
    }
    }
  else{
    if(df[["adj p"]][length(df[["adj p"]])] < 0.05){
      if(df[["Estimate"]][length(df[["Estimate"]])] > 0){
        return("sign_Incr")
      } else{
        return("sign_Decr")
      }
    } else{
      return("Nonsignificant")
  }
  }
}

significances_Spiked <- purrr::map2(results_Spiked, Williams_Spiked, significances) |>
  dplyr::bind_rows(.id = "CompoundLab") |> 
  tidyr::pivot_longer(
    cols = 1:44,  
    names_to = "CompoundLab",        
    values_to = "Spiked_sign"              
  ) |> 
    dplyr::mutate(CompoundLab = stringr::str_remove(CompoundLab,"_Spiked"))


significances_Unspiked <- purrr::map2(results_Unspiked, Williams_Unspiked, significances) |> 
  dplyr::bind_rows(.id = "CompoundLab") |> 
  tidyr::pivot_longer(
    cols = 1:44,  
    names_to = "CompoundLab",        
    values_to = "Unspiked_sign"              
  ) |> 
    dplyr::mutate(CompoundLab = stringr::str_remove(CompoundLab,"_Unspiked"))


desired_order <- c(
  "AmantadineWATCHFR", "AmantadineCEFAS", "AmantadineFIWI", "AmantadineFRAUNH", "AmantidineIDEA",
  "LDopaWATCHFR", "ArabinoseWATCHFR", "CefuroximeWATCHFR", "CefuroximeCEFAS", "CefuroximeFIWI",
  "CefuroximeFRAUNH", "CefuroximeIDEA", "CromolynWATCHFR", "CromolynCEFAS", "CromolynFIWI",
  "CromolynIDEA", "AnastrazoleWATCHFR", "AnastrozoleCEFAS", "AnastrozoleFIWI", "AnastrozoleIDEA",
  "ProchlorazWATCHFR", "DutasterideWATCHFR", "CyproteroneWATCHFR", "FenitrothionWATCHFR",
  "FenitrothionCEFAS", "FenitrothionFIWI", "FenitrothionIDEA", "EstradiolWATCHFR", "mDHTWATCHFR",
  "mDHTCEFAS", "mDHTFIWI", "mDHTFRAUNH", "mDHTIDEA", "TrenboloneWATCHFR", "LinuronWATCHFR",
  "LinuronCEFAS", "LinuronFIWI", "LinuronFRAUNH", "LinuronIDEA", "VinclozolinWATCHFR",
  "VinclozolinCEFAS", "VinclozolinFIWI", "VinclozolinFRAUNH", "VinclozolinIDEA"
)




Result_Radar_valid <- dplyr::full_join(significances_Spiked, significances_Unspiked) |>
    dplyr::mutate(`Known Mode of Action` = 
    dplyr::case_when(
        stringr::str_detect(CompoundLab, "Amantadine|Amantidine|LDopa|Arabinose|Cefuroxime|Cromolyn")~ "no EAS MoA",
        stringr::str_detect(CompoundLab, "Anastrozole|Anastrazole") ~ "steroidogenesis (anti-E): aromatase inhibitor (competitive)",
        stringr::str_detect(CompoundLab, "Prochloraz") ~ "steroidogenesis (anti-E): aromatase inhibitor (transcriptional)",
        stringr::str_detect(CompoundLab, "Dutasteride") ~ "steroidogenesis: 5alpha reductase inhibitor; AR antagonist (high conc.)",
        stringr::str_detect(CompoundLab, "Cyproterone|Fenitrothion")  ~ "anti-A: AR antagonist",
        stringr::str_detect(CompoundLab, "Estradiol") ~ "E: ER agonist; steroidogenesis (anti-A): aromatase inducer",
        stringr::str_detect(CompoundLab, "mDHT|Trenbolone") ~ "A: non-aromatizable AR agonist",
        stringr::str_detect(CompoundLab, "Linuron") ~ "anti-A: AR antagonist; anti-E",
        stringr::str_detect(CompoundLab, "Vinclozolin") ~ "anti-A: AR antagonist; E: weak ER agonist"
    ),
    CompoundLab = factor(CompoundLab, levels = desired_order)) |>
        dplyr::arrange(CompoundLab) |> 
        dplyr::mutate(`Final Conclusion` = c('A(S)-inactive', 'A(S)-inactive', 'A(S)-inactive', 'A(S)-inactive', 'A(S)-inactive', 'A(S)-inactive','active', 'A(S)-inactive', 'A(S)-inactive', 'A(S)-inactive', 'A(S)-inactive', 'A(S)-inactive','A(S)-inactive', 'A(S)-inactive', 'A(S)-inactive', 'A(S)-inactive', 'S-active', 'S-active','S-active', 'inactive', 'inactive', 'A-active', 'A-active', 'A-active', 'A-active', 'A-active', 'A-ative','S-active', 'A-active', 'A-active', 'A-active', 'A-active', 'A-active', 'A-active', 'A-active', 'A-active''inactive', 'inactive', 'A-active', 'A-active', 'A-active', 'A-active', 'inactive', 'A-active')) |> 
    dplyr::relocate(c("CompoundLab", "Known Mode of Action", "Spiked_sign", "Unspiked_sign", "Final conclusion"))
        

gt::gt(Result_Radar_valid)

