---
title: "Software validation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Software validation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


## Reproducible environment

In case you want to make sure that you are creating the exact same output as presented in the publication, you need to create the reproducible environment using `rix` and `nix`. Using `rix`, you can run the following function to create your nix script. The nix script is however already in the vignettes folder within `xeredar` available as default.nix file.

```{r eval = FALSE}
rix::rix(
  r_ver = "4.4.0",
  r_pkgs = c("tidyverse", "tidyr", "purrr", "furrr", "gt","easystats","rlang", "DHARMa", "devtools", "patchwork", "rmarkdown"),
  system_pkgs = NULL,
  tex_pkgs = c("amsmath"),
  git_pkgs = list(package_name = "xeredar",
                  repo_url = "https://github.com/basf/xeredar",
                  commit = "2b85763"),
  ide = "rstudio",
  project_path = "./vignettes",
  overwrite = TRUE
)

```

Once the default.nix file was created in the path you specified in the command above (`project_path`) you need to type `nix-build` into the terminal. Once the shell was built you can open it via `nix-shell`. You can open Rstudio by first running this line in the terminal: `export QT_XCB_GL_INTEGRATION=none` and then running `rstudio`. Rstudio will open and you will have access to all files located at the path you specified in the `rix` function. You can run the script as usual. Through this approach, the code and results are reproducible.

The following sections contain the code that needs to be executed to reproduce the analysis.


## Code for validation

In order to validate the package, we will analyze all 36 XETA ring test studies and produce a table that can be compared with the table in the validation report. The validation studies are available in the `xeredar` package. In case the produced table is in line with the table description, then the package is doing exactly what it is supposed to do and analyzed all 36 XETA ring test studies correctly.

```{r warning=FALSE, message=FALSE, fig.show='hide'}



out <-  !stringr::str_detect(names(xeredar::valid_data_xeta), "_combined$" )
list_dfs <- xeredar::valid_data_xeta[out]
data_prep_trim <- function(df){
  df <- df |> dplyr::filter(!is.na(Fluor))
  dat_trim <- xeredar::trim(df)$dat_trim
  results <- xeredar::data_prep(dat_trim, row= FALSE, trim=FALSE, outlier = FALSE)
  return(results)
}
results <- purrr::map(list_dfs, data_prep_trim,  .progress = TRUE)
worked <- !purrr::map_vec(results, is.list)

difficult_dfs <- list_dfs[worked]

changing_alpha <- c(c(1,5) %o% 10^-(3:17), 0)

decreasing_alpha <- function(df, vector){
  for(i in vector){
    df <- df |> dplyr::filter(!is.na(Fluor))
    dat_trim <- xeredar::trim(df)$dat_trim
    result <- xeredar::data_prep(df, row=FALSE, alpha = i, trim=FALSE, outlier = FALSE)
    if(is.list(result)){
      break
    } else{
      next
    }
  }
  result[["alpha"]] <- i
  return(result)
}


result_difficult_dfs <- purrr::map(.x=difficult_dfs, .f= decreasing_alpha, vector = changing_alpha)
results_alpha_decrease <- results[purrr::map_vec(results, is.list)]

results <- append(results_alpha_decrease, result_difficult_dfs)
check_positive <- function(df){
  if(df[["Monotonicity Test"]]$`Pr(>|t|)`[1] >= 0.01 &
     df[["Monotonicity Test"]]$`Pr(>|t|)`[2] <= 0.01 ) {
    return(df[["Dunnetts"]])
  } else if (length(unique(df$WilliamsIncrease$Y.Tilde)) - 1 > floor(nrow(df$WilliamsIncrease)/2) | 
             length(unique(df$WilliamsDecrease$Y.Tilde)) - 1 > floor(nrow(df$WilliamsDecrease)/2)) { 
    return(list(WilliamsIncrease = df[["WilliamsIncrease"]], WilliamsDecrease = df[["WilliamsDecrease"]])) 
  } else{
    return(df[["Dunnetts"]])
    } 
}

results_2 <- purrr::map(results, check_positive)
Williams <- !purrr::map_vec(results_2, is.data.frame)

significances <- function(df, Williams= FALSE){
  if(Williams) {
    if(length(unique(df$WilliamsIncrease$Y.Tilde)) - 1 > 
       floor(nrow(df$WilliamsIncrease)/2)){
      TRUE %in% df[["WilliamsIncrease"]]$Sign
    } else{
      TRUE %in% df[["WilliamsDecrease"]]$Sign
    }
  } else{
    df[["adj p"]][length(df[["adj p"]])] < 0.05
  }
}

prelim_thyroid_active <- purrr::map2_vec(results_2, Williams, significances)

country_substance_spiked <- names(results_2) |>
  stringr::str_split("\\_") |>
  as.data.frame() |> t() |>
  dplyr::as_tibble() |>
  dplyr::rename("substance" = "V1", "country" = "V2", "spiked" = "V3") |>
  dplyr::mutate(Sign = prelim_thyroid_active)


increase_h12 <- results |> purrr::map_vec(function(x) {
  (x[["SummaryDF"]]$Mean[nrow(x[["SummaryDF"]])] / x[["SummaryDF"]]$Mean[1])-1 
  })


thyroid_active <- country_substance_spiked |> dplyr::mutate(Increase = increase_h12,
                                                            thyroid_active = ifelse(Increase >= 0.12 & Sign == TRUE & spiked == "unspiked" |
                                                                                      abs(Increase) >= 0.12 & Sign == TRUE & spiked == "spiked" , "Thyroid active", "Thyroid inactive")) |> 
  dplyr::select(!Sign, !Increase) |> 
  dplyr::summarise(thyroid_active = ifelse("Thyroid active" %in% thyroid_active, "Thyroid active", "Thyroid inactive"), .by= c(substance, country)) |> 
  dplyr::mutate(
    country = ifelse(country== "japan" & substance %in% c("e2", "nh3") |
                       country == "japan2", "japan lab2", country)) |>
  tidyr::pivot_wider(names_from = country, values_from = thyroid_active) |>
  dplyr::arrange(match(substance, c("t3", "ptu", "t4", "triac", "cefuroxime","linuron", "nh3", "testosterone","e2", "abamectine", "acetone", "isophorone", "metholmyl"))) |> 
  dplyr::mutate(`Expected classification`= c("Thyroid active", "Thyroid active", "Thyroid active", "Thyroid active", "Thyroid inactive", "Thyroid active", "Thyroid active", "Unclear", "Thyroid inactive", "Thyroid inactive","Thyroid inactive","Thyroid inactive","Thyroid inactive"))|>
  dplyr::relocate(c(substance,`Expected classification`, france,japan, usa, `japan lab2`, belgium, portugal ))  |> 
  dplyr::rename(USA = usa)
```

```{r}
gt::gt(thyroid_active, rowname_col = "substance") |>
  gt::tab_options(column_labels.text_transform = "capitalize",
              stub.text_transform = "uppercase") |>
  gt::data_color(columns= -substance, na_color = "white", palette = c("#c2cd94", "#b4c6e7", "white")) |>
  gt::tab_stubhead(label = "Chemical")|>
  gt::tab_spanner(
    label = "Laboratory",
    columns = c(france, japan, USA, 'japan lab2', belgium, portugal)
  ) |>
    gt::tab_style(
    style = list(
      gt::cell_text(color = "#D9654B", weight= "bold")
    ),
    locations = gt::cells_body(
      columns = c(france, belgium), # not needed if coloring all columns
      rows = 8)
  )  |> 
   gt::tab_header(
    title = "Summary table of results from statistical analysis of all 36 validation studies of the ring test of the XETA test guideline. This table shows the results of the automatic analysis carried out with the data_prep() function of the xeredar R-package. Furthermore, this table summarises the results of the original analysis carried out by John Green using SAS. As both approaches led to exactly the same conclusions, only one table is shown. In case a substance was identified to be Thyroid active, the respective cell is colored green, whereas when a substance was identified to be Thyroid inactive, the cell was colored blue. For the study applying testosterone in the french lab, the writing is colored red, because in the original table in the validation document, this substance was incorrectly classified as Thyroid active due to a data transfer mistake. Furthermore, for the study applying testosterone in the belgium lab, the writing is colored red, as well, because in the validation document, a slip in the column led to incorrectly categorizing the respective results as carried out by the japanese lab2.")


```





