title: "Statistical analysis of XETA, RADAR or REACTIV studies"
output: word_document
params: 
  resultList: "resultList"
  conclusion_summary: "conclusion_summary"
  data : "data"



### Data processing and residual diagnostics

`r params$resultList[["Justify"]]`. The p-value from the Shapiro-test is: `r round(params$resultList[["NormalityTest"]]$p.value,3)` and from the Levene test: `r round(params$resultList[["LeveneTest"]][2],3)`. Treatment effects varied from `r round(min(params$resultList$Dunnetts[,6]),2)`% at treatment concentration of `r janitor::clean_names(params$resultList$Dunnetts) |> filter('% Incr'== min('% Incr')) |> rownames() |> stringr::str_extract("\\d+.\\d+|\\d+")` and `r round(max(params$resultList$Dunnetts[,6]),2)`% at treatment concentration of `r janitor::clean_names(params$resultList$Dunnetts) |> filter('% Incr'== max('% Incr')) |> rownames()  |> stringr::str_extract("\\d+.\\d+|\\d+")`.



```{r echo = FALSE, warning = FALSE}
dat <- params$data |>
      dplyr::mutate(
        Removed = !paste(
        params$data$Replicate,
        params$data$Conc,
        params$data$Fluor) 
         %in%
      paste(
        params$resultList[["ProcessedData"]]$Replicate,
        params$resultList[["ProcessedData"]]$Conc,
        params$resultList[["ProcessedData"]]$Fluor)) |>
          gt::gt() |> 
          gt::tab_header(title= "", 
          subtitle= "Uploaded data. The 
          column Removed indicates which measurements
          were removed for further analysis due to potentially required
          trimming or outlier removal for fulfilling the ANOVA 
          assumptions of normally distributed residuals and homogeneous
          variances among treatment groups.")

dat

```


```{r echo = FALSE, warning = FALSE, dpi=600}
norm <- plot(performance::check_normality(params$resultList[["MixedAnova"]]), type = "density")+
    ggplot2::theme(text = ggplot2::element_text(size = 6),
     plot.title = ggplot2::element_text(size = 9), 
     axis.title = ggplot2::element_text(size = 7), 
     axis.text= ggplot2::element_text(size = 5))
hetero <-  plot(performance::check_heteroscedasticity(params$resultList[["MixedAnova"]]))+
    ggplot2::theme(text = ggplot2::element_text(size = 6),
     plot.title = ggplot2::element_text(size = 9),
      axis.title = ggplot2::element_text(size = 7), 
      axis.text= ggplot2::element_text(size = 5))

ggpubr::ggarrange(norm, hetero, ncol = 2)

```

### Results

```{r, warning=FALSE, echo=FALSE, message= FALSE, dpi=600}

params$resultList[["BoxPlots"]]

```


```{r summary of params$resultList processed data, echo= FALSE, warning = FALSE}


params$resultList[["SummaryDF"]] |>
      dplyr::mutate_if(is.numeric, round, 2) |>  
      gt::gt() |> 
      gt::tab_header(title= "Data summary per treatment group.",subtitle="
      The table displays the number of observations (N),  the mean, the standard 
      deviation and the Coefficient of variation
      per treatment. Conc refers to the applied 
      treatment concentration.")
```

```{r Dunnetts test results params$resultList, warning = FALSE, echo= FALSE, label= tbl-DunnettsTest_params$resultList}


params$resultList[["Dunnetts"]] |>
      tibble::rownames_to_column("Contrasts") |>
      dplyr::mutate_if(is.numeric, round, 2) |>
      gt::gt() |>
      gt::tab_header(title="Result table of the dunnett's
       two-sided  many-to-one contrast test result.", subtitle = "
       The table displays the estimated difference (Estimate) and 
       standard error (SE) between the control and the respective
       treatment group. The Kenward-Rogers degrees of freedom (df),
       the t-value and the corresponding adjusted p-value (p)
        are also displayed.")
```

```{r Decreasing Williams test result params$resultList, warning = FALSE, echo = FALSE, label = }

params$resultList[["WilliamsIncrease"]] |>
      dplyr::mutate_if(is.numeric, round, 2) |> 
      gt::gt() |>
      gt::tab_header(title="Result table of the increasing Williams trend test", subtitle ="
      Y.Tilde: The amalgamated means for the respective treatment level;  Y0: The control mean; 
      DIFF Incr: Difference between Y.Tilde and Y0; SE_DIFF: The standard error of the difference; 
      DF: Degrees of freedom; WILL Incr: The value of the increasing Williams test statistic;
      crit Val: The critical value for the increasing Williams test statistic at alpha = 0.05; 
      Sign: In case the value of the increasing Williams test statistic is greater than the critical value, 
      the result is significant (Sign = TRUE), otherwise not significant (Sign = FALSE); % Incr: The %
      increase in the treatment groups in comparison to the control groups.")

```

```{r Increasing Williams test result params$resultList, echo = FALSE, warning = FALSE, label = tbl-WilliamsIncrease_params$resultList}


params$resultList[["WilliamsDecrease"]] |>
      dplyr::mutate_if(is.numeric, round, 2) |>
      gt::gt() |>
      gt::tab_header(title="Result table of the decreasing Williams trend test", subtitle = "
      Y.Tilde: The amalgamated means for the respective treatment level; Y0: The control mean; 
      DIFF Incr: Difference between Y.Tilde and Y0; SE_DIFF: The standard error of the difference; 
      DF: Degrees of freedom; WILL Decr: The value of the decreasing Williams test statistic;
      crit Val: The critical value for the decreasing Williams test statistic at alpha = 0.05; 
      Sign: In case the value of the decreasing Williams test statistic is greater than the critical value, 
      the result is significant (Sign = TRUE), otherwise not significant (Sign = FALSE);% Incr: The %
      increase in the treatment groups in comparison to the control groups. When a decrease is observed
      this value turns negative.")
```

```{r Monotonicity test results params$resultList, echo = FALSE, warning= FALSE, label = tbl-monoTest_params$resultList}
params$resultList[["Monotonicity Test"]] |>
      dplyr::mutate_if(is.numeric, round, 2) |>
      gt::gt() |>
      gt::tab_header(title= "Monotonicity test result 
        specifying the significance of linear and quadratic contrasts 
        on the rank transformed data", subtitle= ". The test results together with 
        the information about the amalgamation of treatment means 
        available in the increasing and decreasing Williams test results, 
        as well as the summary output of the mixed ANOVA model which also 
        provides information about linear and quadratic trends in the data,
        can additionally serve as support to decide about the presence and absence 
        of a monototonic treatment response reslationship.")


```




# Conclusion


```{r final table resultList, warning = FALSE, echo= FALSE, label= tbl-final_table_Unparams$resultList}


params$conclusion_summary |> 
        gt::gt() |> 
      gt::tab_header(title= " ",subtitle="Conclusion table summarizing the percent effect per
       Treatment and the results of the post-hoc tests. The table displays the
       % fluorescence induction in respect to the control. The respective 
       post-hoc test should be selected based on on the dose-response trend in 
       the corresponding data. The monotonicity test, the information about amalgamation of
       treatment means in the Williams test and the summary table of the mixed ANOVA model 
       can be a support for deciding about the presence or absence of a dose-response trend. 
       (ns not significant,. marginally significant 
       ( 0.05 < p > 0.1) ,* significant 
       (p &le; 0.05), n.d. not determined)")

```


\newpage

## R script to produce elements in report

```{r}
#| eval: false
rm(list=ls())

instal <- function(package){
  if(!is.element(package, installed.packages()[,1])){{install.packages(package)}}
}
# StatCharrms is not available on CRAN anymore. 
packages <- c("ARTool","broom","car","dplyr","emmeans","ggplot2","lme4","magrittr","MASS","multcomp","psych","rlang","tidyr", "kableExtra","readxl","knitr","xetaana","stringr")
sapply(packages, instal)
sapply(packages, FUN = function(X){do.call("require", list(X))})

dat <- as.data.frame(read_excel('/Data.xls',sheet = 1, col_names = TRUE))
dat$Replicate <- as.factor(dat$Replicate)
dat$Conc <- factor(dat$Conc, ordered=TRUE, levels=sort(unique(dat$Conc)))
dat$Treatment <- as.factor(dat$Treatment)
dat$Fluor <- as.numeric(dat$Fluor)
dat$Row <- as.factor(dat$Row)

resultList <- xetaana::data_prep(dat, row = TRUE, dagostino = FALSE,
                      trimming = TRUE, outlier = TRUE, boxcox = TRUE,
                      artc = FALSE, alpha = 0.01)


kable(resultList[["SummaryDF"]], digits=2)
kable(resultList[["Dunnetts"]], digits=2)
kable(resultList[["WilliamsDecrease"]], digits=2)
kable(resultList [["WilliamsIncrease"]], digits=2)
kable(resultList[["Monotonicity Test"]],digits=2)


qqnorm(resid(resultList[["MixedAnova"]]), main="",cex=2,cex.lab=1.5, cex.axis=1)
qqline(resid(resultList[["MixedAnova"]]), col = "red", lwd=3, lty=2)           
plot(resid(resultList[["MixedAnova"]])~fitted(params$resultList[["MixedAnova"]]),main="", xlab ="Fitted values",ylab="Pearson residuals",cex=2, cex.lab=1.5, cex.axis=1 )
abline(h = 0, col="blue", lwd=3, lty=2)

resultList[["BoxPlots"]] + theme(text = element_text(size = 14)) 

conclusion_summary <- summary(xeredar::data_summary(resultList))
kable(conclusion_summary)
```
